---
Title: "Lab 5 Student Evaluations of Teaching"
Author: "Bianca Rodas"
---

In this lab, we will be using the `dplyr` package to explore student evaluations
of teaching data. 

**You are expected to use functions from `dplyr` to do your data manipulation!**

# Part 1: GitHub Workflow

Now that you have the Lab 5 repository cloned, you need to make sure you can 
successfully push to GitHub. To do this you need to:

-   Open the `lab-5-student.qmd` file (in the lower right hand corner). 
-   Change the `author` line at the top of the document (in the YAML) to your
name. 
-   Save your file either by clicking on the blue floppy disk or with a shortcut
(command / control + s). 
-   Click the "Git" tab in upper right pane
-   Check the "Staged" box for the `lab-5-student.qmd` file (the file you
changed)
-   Click "Commit"
-   In the box that opens, type a message in "Commit message", such as
"Added my name".
-   Click "Commit".
-   Click the green "Push" button to send your local changes to GitHub.

RStudio will display something like:

```         
>>> /usr/bin/git push origin HEAD:refs/heads/main
To https://github.com/atheobold/introduction-to-quarto-allison-theobold.git
   3a2171f..6d58539  HEAD -> main
```

Now you are ready to go! Remember, as you are going through the lab I would 
strongly recommend rendering your HTML and committing your after **every** 
question! 


# Part 2: Some Words of Advice

Part of learning to program is learning from a variety of resources. Thus, I
expect you will use resources that you find on the internet. There is, however,
an important balance between copying someone else's code and *using their code to learn*.  

Therefore, if you use external resources, I want to know about it.

-   If you used Google, you are expected to "inform" me of any resources you
used by **pasting the link to the resource in a code comment next to where you used that resource**.

-   If you used ChatGPT, you are expected to "inform" me of the assistance you
received by (1) indicating somewhere in the problem that you used ChatGPT 
(e.g., below the question prompt or as a code comment), and (2) downloading and
including the `.txt` file containing your **entire** conversation with ChatGPT.

Additionally, you are permitted and encouraged to work with your peers as you
complete lab assignments, but **you are expected to do your own work**. Copying
from each other is cheating, and letting people copy from you is also cheating.
Please don't do either of those things.

## Setting Up Your Code Chunks

-   The first chunk of this Quarto document should be used to *declare your libraries* (probably only `tidyverse` for now).
-   The second chunk of your Quarto document should be to *load in your data*.

## Save Regularly, Render Often

-   Be sure to **save** your work regularly.
-   Be sure to **render** your file every so often, to check for errors and make
sure it looks nice.
    -   Make sure your Quarto document does not contain `View(dataset)` or `install.packages("package")`, both of these will prevent rendering.
    -   Check your Quarto document for occasions when you looked at the data by
    typing the name of the data frame. Leaving these in means the whole dataset
    will print out and this looks unprofessional. **Remove these!**
    -   If all else fails, you can set your execution options to `error: true`,
    which will allow the file to render even if errors are present.


# Part 3: Let's Start Working with the Data!

## The Data

The `teacher_evals` dataset contains student evaluations of reaching (SET)
collected from students at a University in Poland. There are SET surveys from 
students in all fields and all levels of study offered by the university.

The SET questionnaire that every student at this university completes is as
follows:

> Evaluation survey of the teaching staff of University of Poland. Please
> complete the following evaluation form, which aims to assess the lecturer’s
> performance. Only one answer should be indicated for each question. The
> answers are coded in the following way: 5 - I strongly agree; 4 - I agree;
> 3 - Neutral; 2 - I don’t agree; 1 - I strongly don’t agree.
>
> Question 1: I learned a lot during the course.
>
> Question 2: I think that the knowledge acquired during the course is very
> useful.
>
> Question 3: The professor used activities to make the class more engaging.
>
> Question 4: If it was possible, I would enroll for a course conducted by this
> lecturer again.
>
> Question 5: The classes started on time.
>
> Question 6: The lecturer always used time efficiently.
>
> Question 7: The lecturer delivered the class content in an understandable and
> efficient way.
>
> Question 8: The lecturer was available when we had doubts.
>
> Question 9. The lecturer treated all students equally regardless of their
> race, background and ethnicity.

These data are from the end of the winter semester of the 2020-2021 academic
year. In the period of data collection, all university classes were entirely
online amid the COVID-19 pandemic. While expected learning outcomes were not
changed, the online mode of study could have affected grading policies and could
have implications for data.

**Average SET scores** were combined with many other variables, including:

1.  **characteristics of the teacher** (degree, seniority, gender, SET scores in
the past 6 semesters).
2.  **characteristics of the course** (time of day, day of the week, course
type, course breadth, class duration, class size).
3.  **percentage of students providing SET feedback.**
4.  **course grades** (mean, standard deviation, percentage failed for the
current course and previous 6 semesters).

This rich dataset allows us to **investigate many of the biases in student evaluations of teaching** that have been reported in the literature and to formulate new
hypotheses.

Before tackling the problems below, study the description of each variable
included in the `teacher_evals_codebook.pdf`.

**1. Load the appropriate R packages for your analysis.**

```{r}
#| label: setup
# code chunk for loading packages
library(tidyverse)
library(dplyr)
```

**2. Load in the `teacher_evals` data.** 

```{r}
#| label: load-data
# code chunk for importing the data
eval <- read_csv("data-raw/teacher_evals.csv")
```

### Data Inspection + Summary

**3. Provide a brief overview (~4 sentences) of the dataset.**

```{r}
#| label: explore-data
# you may want to use code to answer this question
#glimpse(eval)
summary(eval)
```
>The data set contains 8,015 rows and 22 columns/ variables. All but five variables are read as doubles, while the remaining are read in as characters. We have information that range from the teacher's characteristics, the course characteristics, the students feedback, and their course grades.  


**4. What is the unit of observation (i.e. a single row in the dataset) identified by?**

```{r}
#| label: row-identification
# you may want to use code to answer this question
head(eval)
```

> The unit of observation are the answers to the questions per course and teacher.  

**5. Use _one_ `dplyr` pipeline to clean the data by:**

- **renaming the `gender` variable `sex`**
- **removing all courses with fewer than 10 respondents**
- **changing data types in whichever way you see fit (e.g., is the instructor ID really a numeric data type?)**
- **only keeping the columns we will use -- `course_id`, `teacher_id`, `question_no`, `no_participants`, `resp_share`, `SET_score_avg`, `percent_failed_cur`, `academic_degree`, `seniority`, and `sex`**

**Assign your cleaned data to a new variable named `teacher_evals_clean` –- use these data going forward. Save the data as `teacher_evals_clean.csv` in the `data-clean` folder.**

```{r}
#| label: data-cleaning
# code chunk for Q4
teacher_evals_clean <- eval |> 
  rename(sex=gender) |> 
  filter(no_participants >= 10) |> 
  mutate(
    teacher_id = as.numeric(teacher_id), 
    course_id = as.character(course_id),
    question_no = as.numeric(question_no),
    academic_degree = as.factor(academic_degree),
    seniority = as.factor(seniority),
    sex = as.character(sex)) |> 
  select(course_id, teacher_id, question_no, no_participants, 
  resp_share, SET_score_avg, percent_failed_cur, academic_degree, 
  seniority, sex)
  
write.csv(teacher_evals_clean, file="data-clean/teacher_evals_clean.csv", row.names=FALSE)
```

**6. How many unique instructors and unique courses are present in the cleaned dataset?**

```{r}
#| label: unique-courses
# code chunk for Q5
teacher_evals_clean |> 
  summarise(n_distinct(teacher_id))

teacher_evals_clean |> 
  summarise(n_distinct(course_id))
```
> There are 297 unique instructors and 939 unique courses.  


**7. One teacher-course combination has some missing values, coded as `NA`. Which instructor has these missing values? Which course? What variable are the missing values in?**

```{r}
#| label: uncovering-missing-values
# code chunk for Q6

missing_info <- teacher_evals_clean |> 
  mutate(row_id = row_number()) |>
  select(where(is.numeric), teacher_id, course_id, row_id) |>
  pivot_longer(cols = -c(row_id, teacher_id, course_id),
               names_to = "variable",
               values_to = "value") |> 
  filter(is.na(value)) |> 
  select(teacher_id, 
         course_id,
         variable) |> 
  distinct()

missing_info
```
> The teacher-course combination with missing values is teacher id, 56347 ,and course id, PAB3SE004PA. The variable with the missing values is the percent_failed_cur.   

**8. What are the demographics of the instructors in this study? Investigate the variables `academic_degree`, `seniority`, and `sex` and summarize your findings in ~3 complete sentences.**

```{r}
#| label: exploring-demographics-of-instructors
# code chunk for Q7

instructor_demographics <- teacher_evals_clean |>
  select(academic_degree, seniority, sex) |>
  distinct() |>
  summarise(
    total_instructors = n(),
    degree = list(table(academic_degree)),
    seniority = list(table(seniority)),
    sex = list(table(sex)))


instructor_demographics$degree
instructor_demographics$seniority
instructor_demographics$sex
```
> For academic degrees, there are 22 doctorates, 19 masters, 6 n/a, and 7 professors degrees. For seniority, there is ranging between 1-11 years, with majority around 4 and 5 years of teaching at the school. As for sex, there are 23 females and 31 male teachers.  

**9. Each course seems to have used a different subset of the nine evaluation questions. How many teacher-course combinations asked all nine questions?**

```{r}
#| label: teacher-course-asked-every-question
# code chunk for Q8
teacher_evals_clean |> 
  select(resp_share = 1) 

tcaeq <- teacher_evals_clean |>
  group_by(teacher_id, course_id) |> 
  summarise(n_questions = n_distinct(question_no), .groups = "drop") |> 
  filter(n_questions == 9)|>
  count(n_questions) 

tcaeq
```
> 49 teacher-course combinations asked all nine questions


## Rate my Professor

**10. Which instructors had the highest and lowest average rating for Question 1 (I learnt a lot during the course.) across all their courses?**

```{r}
#| label: question-1-high-low
# code chunk for Q9
summary(teacher_evals_clean$question_no)

q1_ratings <- teacher_evals_clean |> 
  filter(question_no == 901) |> 
  group_by(teacher_id) |> 
  summarise(avg_q1 = mean(SET_score_avg, na.rm = TRUE), .groups = "drop")


highest_avg <- q1_ratings |> 
  filter(avg_q1 == max(avg_q1))

lowest_avg  <- q1_ratings |> 
  filter(avg_q1 == min(avg_q1))

highest_avg
lowest_avg
```
>The instructor with the highest rating is teacher_id 54201, and the lowest rating is teacher_id 100132. 

**11. Which instructors with one year of experience had the highest and lowest average percentage of students failing in the current semester across all their courses?**

```{r}
#| label: one-year-experience-failing-students
# code chunk for Q10
fail_rate_by_new_teacher <- teacher_evals_clean |> 
  filter(seniority == 1) |> 
  group_by(teacher_id) |> 
  summarise(avg_percent_failed = mean(percent_failed_cur, na.rm = TRUE), .groups = "drop")

#Get teachers with highest and lowest average failure rates
highest_fail_rate <- fail_rate_by_new_teacher |> 
  filter(avg_percent_failed == max(avg_percent_failed, na.rm = TRUE))

lowest_fail_rate <- fail_rate_by_new_teacher |> 
  filter(avg_percent_failed == min(avg_percent_failed, na.rm = TRUE))

highest_fail_rate
lowest_fail_rate
```
> The instructor with the highest fail rate with one year of experience is the teacher_id 106692. The instructor with the lowest are teacher_id's 86222, 98650, 98651, 102379, 103092, 106126. 

**12. Which female instructors with either a doctorate or professional degree had the highest and lowest average percent of students responding to the evaluation across all their courses?**

```{r}
#| label: female-instructor-student-response
# code chunk for Q11

female_doctorate_pro <- teacher_evals_clean |> 
  filter(sex == "female",
    academic_degree %in% c("dr", "prof")) |> 
  group_by(teacher_id) |> 
  summarise(avg_resp_share = mean(resp_share, na.rm = TRUE), .groups = "drop")

highest_resp <- female_doctorate_pro  |> 
  filter(avg_resp_share == max(avg_resp_share, na.rm = TRUE))

lowest_resp <- female_doctorate_pro |> 
  filter(avg_resp_share == min(avg_resp_share, na.rm = TRUE))

highest_resp
lowest_resp
```

> The highest average response is teacher_id 101508. The lowest response is 59338.  



